\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Research Paper: Cyberattack Case Study \\CIS 3360-Security in Computing\\Fall 2025\\
{\footnotesize Instructor: Jie Lin, Ph.D.}
\thanks{}
}

\author{{ Derek Oliveira}\\
\IEEEauthorblockA{\textit{department of engeneering and computer science} \\
Orlando, United States of America \\
de937084@ucf.edu}
}

\maketitle

\begin{abstract}
The SolarWinds cyber breach, disclosed in December 2020, stands as one of the most significant cyber-espionage
incidents of the modern era. Through a sophisticated supply-chain compromise of SolarWinds' Orion platform, 
the state sponsored hacker groupd NOBELIUM gained persistent and covert access to U.S. government agencies,
corporations, and critical service providers. I examine the historical context, motives, and technical execution
of the attack that gave persistent anonymous access to SolarWinds' build environment, dubbed SUNBURST, demonstrating how the breach not 
only infiltrated high level networks but also undermined global trust in software ecosystems. By analyzing
NOBELIUM's tactics and the far reaching consequences of the intrusion, I argue that the SolarWinds incident
represents a wakeup call moment in cybersecurity. It highlights the vulnerability of supply chains, exposes
the limitations of traditional defense mechanisms, and guided us to new industry standards that shape the cybersecurity
landscape today. 
\end{abstract}

\begin{IEEEkeywords}
SUNBURST, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
On December 13, 2020, Solarwinds, A Texas based technology company that develops network monitoring and IT 
infrastructure management software, announced what would become the most consequential cyber intrusion in history.
The breach focused on exploiting their Orion software, SolarWinds' main network monitoring platform used by U.S. 
government agencies, defense contractors, Fortune 500 coorporations, and critical national service providers. 
The attackers, latter dubbed NOBELIUM by microsoft's Threat Intelligence center (MSTIC) and discovered to be
state sponsored attacker group backed by Russia, pursued espionage rather than economic gains.
SolarWinds itself was not the ultimate target, thou. Instead, its compromised software supply chain became the conveyor belt for malicious 
code that infiltrated networks of its customer base undetected for over a year. There where many exploits deployed in this attack, but for the 
scope of this research I will focus on the "SUNBURST" exploit.

By the time the intrusion was detected, NOBELIUM had already garanteed their persistent, undetected access to email servers,
sensitive repositories, and classified communications. What followed was not a simple data breach, but a calculated and methodical 
campaign of cyber-espionage that exploited trust and used it as a mask for malicious intent. The SolarWinds incident was deemed a 
significant cyberattack not only because of its scale and sophistication, but also because of the national security risk it posed and 
the trust it shattered in the global technology leaders that are the watchdogs of our networks.

I argue that the SolarWinds breach represents a turning point in cybersecurity. It revealed the vulnerabilities of software supply 
chains, exposed the weakness of existing detection systems, and forced governments and corporations to rethink the resilience of their 
networks and how they can restructure themselves for reduced impact. To explore this claim, I will first outline the historical 
background and motives of NOBELIUM, then examine the technical methods of intrusion, followed by an analysis of the attack's distinct 
features, its broader geopolitical implications and if the attack's patterns are still in use today. Finally, I will evaluate the lessons 
learned and how this incident reshaped modern approaches to cyber defense.

\section{Background}

Background text here

\section{analysis}

Analysis text here.


\section*{discussion}

Discussion text here.

\section{conclusion}

Conclusion text here.

\section*{citations}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}
About
About us
Careers
Blog
Solutions
For business
For universities
For government
For publishers
Customer stories
Learn